Another stratification

### 2024-07-30 test_boot_short.Rmd
Test for better bootstrap variance

---
title: "Monte Carlo and non-uniform bootstrap"
output: pdf_document
date: "2024-07-29"
---

@@

for (l in 0:((1000/c_dim)-1)) {
  if (l==0){
    x_init <- i*c_dim
  }else if ((3*l+i)*c_dim < (10)*c_dim){
    x_init <- (3*l+i)*c_dim
  }else{
    x_init <- (((3*l+i)*c_dim)%%(c_dim*10))
  }
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    random_points <- rbind(random_points, list(a,b, l))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y", "line")
return(random_points)
}
```

@@

# Monte Carlo alone

10 000 Monte Carlo iterations are appropriated.

```{r}
### Function to automate Monte Carlo method

MonteCarlo <- function(n, B){
  # B: number of iterations for MC

  p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density
  MC_df <- data.frame(useless = rep(0, n))
  MC_est <- data.frame(useless = rep(0, 3), row.names = c("est_nb", "est_vol", "est_ratio"))

  for (b in 1:B) { # All in one "for" loop
    # Sampling over the whole territory
    df_grid <- sampling_grid(i,c_dim)
    a1 <- df_grid$x
    a2 <- df_grid$y
    a3 <- df_grid$line
    # n <- nrow(df_grid) # sample size
    # p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density
    # Variables of interest derived from the sample
    var <- mapply(circle_all, a1, a2)
    MC_b <- data.frame(x = a1, y = a2, w = a3, z = t(var)[,1], t = t(var)[,2])
    MC_df <- cbind(MC_df, MC_b)
    colnames(MC_df)[(5*(b-1)+2):(5*(b-1)+2+4)] <- c(paste("x", b, sep = ""), paste("y", b, sep = ""), paste("l", b, sep = ""), paste("nb", b, sep = ""), paste("vol", b, sep = ""))
    est_b <- t(data.frame(x = sum(MC_df[5*(b-1)+2+3] /p), y = sum(MC_df[5*(b-1)+2+4] /p), z = (sum(MC_df[5*(b-1)+2+4] /p))/(sum(MC_df[5*(b-1)+2+3]/p))))
    MC_est <- cbind(MC_est, est_b)
    colnames(MC_est)[b+1] <- paste("ech", b, sep = "") 
  }
  MC_df$useless <- NULL
  MC_est$useless <- NULL
  MC_var <- c(var(as.vector(t(MC_est[1,]))), var(as.vector(t(MC_est[2,]))), var(as.vector(t(MC_est[3,]))))

  # Confidence intervals
  norm_CI <- data.frame(useless = rep(0, 2))
  perc_CI <- data.frame(useless = rep(0, 2))
  rev_perc_CI <- data.frame(useless = rep(0, 2))
  for (l in 1:3) { # Loop: one estimator each time
    est_init <- as.vector(t(MC_est[l,]))[1]
    sorted_est <- sort(as.vector(t(MC_est[l,])))
    lim_inf <- sorted_est[floor(B*2.5/100)]
    lim_sup <- sorted_est[B - floor(B*2.5/100)]
    perc_CI <- cbind(perc_CI, data.frame(c(lim_inf, lim_sup)))
    rev_perc_CI <- cbind(rev_perc_CI, data.frame(c(est_init*2 - lim_sup, est_init*2 - lim_inf)))
    norm_CI <- cbind(norm_CI, data.frame(c(est_init-qnorm(0.975)*sqrt(MC_var[l]), est_init+qnorm(0.975)*sqrt(MC_var[l]))))
  }
  norm_CI$useless <- NULL
  perc_CI$useless <- NULL
  rev_perc_CI$useless <- NULL
  colnames(norm_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  colnames(perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  colnames(rev_perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  all_CI <- list(Normal = norm_CI, Percentile = perc_CI, `Reverse percentile` = rev_perc_CI)

  return(list(`Local Variables` = MC_df, Estimations = MC_est, Variances = MC_var, `Confidence Intervals` = all_CI))
  # Local Variables: local variables (number and volume) for each point in each sample
  # Estimations: 3 Estimations (number, volume, ratio volume / number) for each sample
  # Variances: Monte Carlo variance for each of the totals / ratio
  # Confidence Intervals: 3 different forms of 95% confidence intervals, for each total / ratio
}
```

@@

  df_boot <- data.frame(useless = rep(0, nrow(df_sample)-10))
  for (b in 1:B) {
    nb <- sample(unique(df_sample[,3]), (length(unique(df_sample[,3]))-1), replace = TRUE)
    # (H-1) lines at random
    sample_b <- data.frame(x = numeric(), y = numeric(), z = numeric(), t = numeric(), w = numeric())
    for (number in nb){
      sample_b <- rbind(sample_b, df_sample[df_sample[,3] == number,])
    }
    df_boot <- cbind(df_boot, sample_b)
    colnames(df_boot)[((b-1)*5+1 + 1):(b*5+1)] <- c(paste("ech", b, colnames(df_sample)[1], sep = ""), paste("ech", b, colnames(df_sample)[2], sep = ""), paste("ech", b, colnames(df_sample)[3], sep = ""), paste("ech", b, colnames(df_sample)[4], sep = ""), paste("ech", b, colnames(df_sample)[5], sep = ""))
  }
  df_boot$useless <- NULL
  return(df_boot)
  # Return a data frame with B samples of as many variables as we want (in this application, 5)
}
```

@@

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Uniform sampling over the whole territory
    df_grid <- sampling_grid(i,c_dim)
    a1 <- df_grid$x
    a2 <- df_grid$y
    a3 <- df_grid$line
    var <- mapply(circle_all, a1, a2)
    df_b <- data.frame(x = a1, y = a2, w = a3, z = t(var)[,1], t = t(var)[,2])
    colnames(df_b) <- c(paste("x.MC", b, sep = ""), paste("y.MC", b, sep = ""), paste("l.MC", b, sep = ""), paste("nb.MC", b, sep = ""), paste("vol.MC", b, sep = ""))
    est_b <- t(data.frame(x = sum(df_b[4] /p), y = sum(df_b[5] /p), z = (sum(df_b[5] /p))/(sum(df_b[4]/p))))

@@

## Monte Carlo alone

res1 <- MonteCarlo(n, 10000)
# Just one circle to begin with

```{r}
cat("Monte Carlo means of estimations: \n - for the number of trees:", mean(as.vector(t(res1$Estimations[1,]))), "\n - for the volume of trees:", mean(as.vector(t(res1$Estimations[2,]))), "\n - for the ratio volume / number of trees:", mean(as.vector(t(res1$Estimations[3,]))))

```{r}
cat("Monte Carlo variances: \n - for the number of trees:", res1$Variances[1], "\n - for the volume of trees:", res1$Variances[2], "\n - for the ratio volume / number of trees:", res1$Variances[3])

```{r}
cat("95% confidence intervals using the normal method:\n - for the number of trees: [", res1$`Confidence Intervals`$Normal[1,1], ";", res1$`Confidence Intervals`$Normal[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$Normal[1,2], ";", res1$`Confidence Intervals`$Normal[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$Normal[1,3], ";", res1$`Confidence Intervals`$Normal[2,3],
    "]")

```{r}
cat("95% confidence intervals using the percentile method:\n - for the number of trees: [", res1$`Confidence Intervals`$Percentile[1,1], ";", res1$`Confidence Intervals`$Percentile[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$Percentile[1,2], ";", res1$`Confidence Intervals`$Percentile[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$Percentile[1,3], ";", res1$`Confidence Intervals`$Percentile[2,3],
    "]")

```{r}
cat("95% confidence intervals using the reverse percentile method:\n - for the number of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,1], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,2], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,3], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,3],
    "]"
    )

@@

```{r}
diff_percent <- function(x,y)100*(max(x,y)-min(x,y))/min(x,y)
cat("Difference between the two estimations of variance: \n - for the number of trees: ", diff_percent(res1$Variances[1], mean(VAR_nb)), "%", "\n - for the volume of trees: ", diff_percent(res1$Variances[2], mean(VAR_vol)), "%\n - for the ratio volume / number of trees: ", diff_percent(res1$Variances[3], mean(VAR_ratio)), "%", sep = "")

>>>>

---
title: "Another stratification"
output: pdf_document
date: "2024-07-29"
---

@@

for (l in 0:((1000/c_dim)-1)) {
  if (l==0){
    x_init <- i*c_dim
  }else if ((3*l+i)*c_dim < (10)*c_dim){
    x_init <- (3*l+i)*c_dim
  }else{
    x_init <- (((3*l+i)*c_dim)%%(c_dim*10))
  }
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    h <- nrow(random_points)%%100 # 100 strata
    random_points <- rbind(random_points, list(a,b, h))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y", "Stratum")
return(random_points)
}
```

@@

@@

  df_boot <- data.frame(useless = rep(0, nrow(df_sample)-10))
  for (b in 1:B) {
    nb <- sample(unique(df_sample[,3]), (length(unique(df_sample[,3]))-1), replace = TRUE)
    # (H-1) strata at random
    sample_b <- data.frame(x = numeric(), y = numeric(), z = numeric(), t = numeric(), w = numeric())
    for (number in nb){
      sample_b <- rbind(sample_b, df_sample[df_sample[,3] == number,])
    }
    df_boot <- cbind(df_boot, sample_b)
    colnames(df_boot)[((b-1)*5+1 + 1):(b*5+1)] <- c(paste("ech", b, colnames(df_sample)[1], sep = ""), paste("ech", b, colnames(df_sample)[2], sep = ""), paste("ech", b, colnames(df_sample)[3], sep = ""), paste("ech", b, colnames(df_sample)[4], sep = ""), paste("ech", b, colnames(df_sample)[5], sep = ""))
  }
  df_boot$useless <- NULL
  return(df_boot)
  # Return a data frame with B samples of as many variables as we want (in this application, 5)
}
```

@@

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Uniform sampling over the whole territory
    df_grid <- sampling_grid(i,c_dim)
    a1 <- df_grid$x
    a2 <- df_grid$y
    a3 <- df_grid$Stratum
    var <- mapply(circle_all, a1, a2)
    df_b <- data.frame(x = a1, y = a2, w = a3, z = t(var)[,1], t = t(var)[,2])
    colnames(df_b) <- c(paste("x.MC", b, sep = ""), paste("y.MC", b, sep = ""), paste("h.MC", b, sep = ""), paste("nb.MC", b, sep = ""), paste("vol.MC", b, sep = ""))
    est_b <- t(data.frame(x = sum(df_b[4] /p), y = sum(df_b[5] /p), z = (sum(df_b[5] /p))/(sum(df_b[4]/p))))

@@

@@

```{r}
diff_percent <- function(x,y)100*(max(x,y)-min(x,y))/min(x,y)
cat("Difference between the two estimations of variance: \n - for the number of trees: ", diff_percent(240516.7, mean(VAR_nb)), "%", "\n - for the volume of trees: ", diff_percent(14084.96, mean(VAR_vol)), "%\n - for the ratio volume / number of trees: ", diff_percent(6.397115e-06, mean(VAR_ratio)), "%", sep = "")
